## This file includes:
## 1. Color code the combined UMAP according to time and condition
## 2. Plot relevant markers in a dot plot 
## 3. Bar plot of cell type composition at each time point 
## 4. Gene expression heatmaps/trends 


# Visualize UMAP annotations 
```{r}
# Color code the combined UMAP according to days 
DimPlot(data, reduction = "umap", group.by=="day")

# Color code the combined UMAP according to conditions 
DimPlot(data, reduction = "umap", group.by=="condition")
```



# Plot relevant markers in a dot plot  
```{r}
# Create a new column in meta data called day_celltypes that contain both day and celltypes information 
combined$day_celltypes = paste(combined$day, combined$celltypes, sep = "_") 
genes = c("SOX2", "TNC", "LIFR", "HOPX", "LHX2", "EOMES", "TOP2A", "MKI67", "TBR1", "BCL11B", "RELN", "SATB2", "CUX2", "GAD1", "DLX5", "DLX6", "RSPO2", "LMX1A", "OTX2", "TTR", "PAX3", "PAX7", "NEAT1", "VEGFA")

# Visualize dot plot 
DotPlot(combined, features = genes, group.by="day_celltypes") + RotatedAxis() 
```


# Cell type composition bar plot 
```{r}
# "#e5c494" 
library(ggplot2)
library(dplyr)
library(RColorBrewer)
custom_colors <- c("#1f78b4", "#33a02c", "#e31a1c", "#ff7f00", "#6a3d9a", 
                   "#b15928", "#a6cee3", "#b2df8a", "#fb9a99", "#fdbf6f", 
                   "#cab2d6", "#ffff99", "#a6d854", "#ffd92f", 
                   "#8da0cb")

metadata <- combined@meta.data

# cell number per sample: 
cell_type_summary <- metadata %>%
  group_by(orig_samples,celltypes) %>%
  summarise(n = n()) %>%
  ungroup()

# cell type per sample: 
cell_type_summary <- metadata %>%
  group_by(celltypes, condition) %>%
  summarise(n = n()) %>%
  ungroup()

# Calculate proportions of each cell type within each time point
cell_type_summary <- cell_type_summary %>%
  group_by(condition) %>%
  mutate(proportion = n / sum(n))

# compare cell type information by conditions 
cell_type_summary <- cell_type_summary %>%
  mutate(condition = factor(condition, levels = c("FGF","ctr","BMP")))

# Visualize cell type summary 
ggplot(cell_type_summary, aes(x = condition, y = proportion, fill = celltypes)) +
  geom_bar(stat = "identity", position = "fill") +  # Stacked barplot with proportions
  theme_minimal() +
  labs(x = "Condition", y = "Proportion", fill = "Cell Type") +
  scale_fill_manual(values = custom_colors) +  # Use a color palette for cell types
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels
```


# Plot gene expression heatmaps 

```{r}
library(readxl)
file_path <- "your_path" # Modify to your own path of file 
your_file <- as.data.frame(read_excel(file_path)) # Read the file that contains all gene names you wish to visualize (as an example, see the Excel sheet sent by Momoko) 
```


```{r}
library(pheatmap)
library(RColorBrewer)

D35_sample = c("YC44","KS108")
D56_sample = c("YC36","YC46")
D97_sample = c("YC34","YC40")

# Create a dictionary mapping shorthand to full cell type names
celltype_dict <- list(
  "ULN" = "ULN",
  "DLN" = "DLN",
  "NN" = "newborn neurons",
  "IP" = "IP",
  "aRGoRG" = "aRG/oRG",
  "CC" = "cycling cells",
  "PPCR" = "PP/CR",
  "aRG" = "aRG"  
)

for (colname in colnames(Momoko_file)) {
  print(colname)
  split_parts <- unlist(strsplit(colname, "_"))
  day <- split_parts[1] 
  celltype <- split_parts[2] 
  if (celltype != "aRG"){
    subdata = subset(combined, day == day & celltypes == celltype_dict[[celltype]])}
  else{
    subdata = subset(combined, day==day & (celltypes == "aRG"|celltypes=="oRG-like cells"))
  }
  gene_ls = your_file[[colname]][!is.na(your_file[[colname]])]
  res = AverageExpression(subdata,features=gene_ls,group.by=c("celltypes","condition","orig_samples")) 
  
  if (day == "D35") {
    sample_id <- D35_sample
  } else if (day == "D56") {
    sample_id <- D56_sample
  } else {
    sample_id <- D97_sample  # Optional: handle other cases if needed
  }
  
  if (celltype != "aRG"){
    data <- res$RNA[, c(paste0(celltype_dict[[celltype]], "_FGF_", sample_id[1], "-", day, "-FGF"),
                        paste0(celltype_dict[[celltype]], "_FGF_", sample_id[2], "-", day, "-FGF"),
                        paste0(celltype_dict[[celltype]], "_ctr_", sample_id[1], "-", day, "-ctr"),
                        paste0(celltype_dict[[celltype]], "_ctr_", sample_id[2], "-", day, "-ctr"),
                        paste0(celltype_dict[[celltype]], "_BMP_", sample_id[1], "-", day, "-BMP"),
                        paste0(celltype_dict[[celltype]], "_BMP_", sample_id[2], "-", day, "-BMP"))]}
  else{
    data <- res$RNA[, c(paste0("oRG-like cells", "_FGF_", sample_id[1], "-", day, "-FGF"),
                        paste0("oRG-like cells", "_FGF_", sample_id[2], "-", day, "-FGF"),
                        paste0(celltype_dict[[celltype]], "_FGF_", sample_id[1], "-", day, "-FGF"), 
                        paste0(celltype_dict[[celltype]], "_FGF_", sample_id[2], "-", day, "-FGF"),
                        paste0(celltype_dict[[celltype]], "_ctr_", sample_id[1], "-", day, "-ctr"),
                        paste0(celltype_dict[[celltype]], "_ctr_", sample_id[2], "-", day, "-ctr"),
                        paste0(celltype_dict[[celltype]], "_BMP_", sample_id[1], "-", day, "-BMP"),
                        paste0(celltype_dict[[celltype]], "_BMP_", sample_id[2], "-", day, "-BMP"))]   
  }
  data = log2(data+1e-5)  
  breaks <- seq(-5, 5, length.out = 100) 
  your_dir = "..." # Your own direction + /figure name 
  pdf(paste0(your_dir, colname, ".pdf"), width = 6, height = length(gene_ls)/5+5)
  my_palette <- colorRampPalette(rev(brewer.pal(n=11, name='RdBu')))(100)
  pheatmap(data, breaks=breaks,cluster_cols = FALSE,cluster_rows=FALSE, color = my_palette) 
  dev.off()
}
```


# Plot gene expression heatmaps (Optional: use this if you wish to work one by one for each sample manually) 
```{r}
D97_ULN = subset(combined, day == "D97" & celltypes == "ULN")
D97_DLN = subset(combined, day == "D97" & celltypes == "DLN")
D97_NN = subset(combined, day == "D97" & celltypes == "newborn neurons")
D97_IP = subset(combined, day == "D97" & celltypes == "IP")
D97_aRGoRG = subset(combined, day == "D97" & celltypes == "aRG/oRG")
D97_CC = subset(combined, day == "D97" & celltypes == "cycling cells")
D56_DLN = subset(combined, day == "D56" & celltypes == "DLN")
D56_NN = subset(combined, day == "D56" & celltypes == "newborn neurons")
D56_IP = subset(combined, day == "D56" & celltypes == "IP")
D56_CC = subset(combined, day == "D56" & celltypes == "cycling cells")
D35_PPCR = subset(combined, day == "D35" & celltypes == "PP/CR")
D35_CC = subset(combined, day == "D35" & celltypes == "cycling cells")
D35_IP = subset(combined, day == "D35" & celltypes == "IP")
D35_aRG = subset(combined, day == "D35" & (celltypes == "aRG"|celltypes=="oRG-like cells"))  
D56_aRG = subset(combined, day == "D56" & (celltypes == "aRG"|celltypes=="oRG-like cells"))  
```


```{r}
gene_ls = c("gene1", "gene2") # Change to genes of your interest
length(gene_ls)
res = AverageExpression(data,features=gene_ls,group.by=c("celltypes","condition","orig_samples")) 
```


```{r}
library(pheatmap)
colnames = colnames(res$RNA)
data <- res$RNA[,c("DLN_FGF_YC34-D97-FGF", "DLN_FGF_YC40-D97-FGF","DLN_ctr_YC34-D97-ctr", "DLN_ctr_YC40-D97-ctr","DLN_BMP_YC34-D97-BMP", "DLN_BMP_YC40-D97-BMP")] # Change to the form: celltype_condition_batch-day-condition 
data = log2(data+1e-5)  
breaks <- seq(-5, 5, length.out = 100) 
pdf("your_dir", width = 6, height = length(gene_ls)/5) # Save figure to your desired directory 
pheatmap(data, breaks=breaks,cluster_cols = FALSE,cluster_rows=FALSE) 
```


# Go terms enrichment analysis
```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(DOSE)
```


```{r}
genes = c("gene1","gene2") # Change into the genes of your interest 

ego <- enrichGO(gene         = genes,
                OrgDb        = org.Hs.eg.db,
                keyType      = "SYMBOL",      
                ont          = "ALL",         # Ontology: BP, CC, MF, or ALL
                pAdjustMethod = "BH",         # p-value adjustment method
                pvalueCutoff  = 0.05,         # p-value cutoff
                qvalueCutoff  = 0.05,         # q-value cutoff
                readable      = FALSE)
# View results
head(ego)
dotplot(ego, showCategory = 10, orderBy = "GeneRatio") + ggtitle("GO Enrichment Dotplot")
```










