## This file includes:
## 1. Preprocessing 
## 2. Clustering
## 3. Subclustering
## 4. Combine different samples 

## Seurat preprocessing and clustering 
```{r}
library(Seurat)
library(harmony)
library(dplyr)
library(patchwork)
library(SeuratData)
data=Read10X(data.dir="../Raw_data/all_well/",gene.column=1)
data=CreateSeuratObject(counts = data, project = "all_well", min.cells = 3, min.features = 200)
```


```{r}
data[["percent.mt"]] <- PercentageFeatureSet(data, pattern = "^MT-") 
data <- subset(data, subset = nFeature_RNA > 500 & nFeature_RNA < 25000 & percent.mt<5) 
data <- NormalizeData(data, normalization.method = "LogNormalize", scale.factor = 10000)
data <- FindVariableFeatures(data, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(data)
data <- ScaleData(data, features = all.genes)
data <- RunPCA(data, features = VariableFeatures(object = data))
data <- FindNeighbors(data, dims = 1:15)
data <- FindClusters(data, resolution = 0.5)
data <- RunUMAP(data, dims = 1:15)
```


```{r}
# Visualize clustering in UMAP 
DimPlot(data, reduction = "umap")
```


## Find markers
```{r}
data.markers <- FindAllMarkers(data, only.pos = TRUE)
data.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1)
```


## Draw feature plots of different genes 
```{r}
FeaturePlot(data, features = c("gene1","gene2")) # Change gene names to your desired gene names 
```


## Assign cell types 
```{r}
new.cluster.ids <- c("cluster1_name","cluster2_name","cluster3_name","cluster4_name") # Change the cluster names to your desired cluster names 
names(new.cluster.ids) <- levels(data)
data <- RenameIdents(data, new.cluster.ids)
```


## Visualize cell type clustering in UMAP 
```{r}
DimPlot(data, reduction = "umap") 
```


## Save data
```{r}
saveRDS(data, file = "../data.rds") 
```


## Subclustering

# Subcluster "Choroid plexus+alpha" into 2 subclusters

```{r}
# Lots of iterations are done to obtain the optimal subclusters 
cluster = subset(data, idents = "CP") # If you wish to sucbluster another cluster, change idents to your desired cluster
cluster <- RunPCA(cluster, features = VariableFeatures(object = cluster))
cluster <- FindNeighbors(cluster, dims = 1:10)
cluster <- FindClusters(cluster, resolution = 0.2)
cluster <- RunUMAP(cluster, dims = 1:10)
DimPlot(cluster, reduction = "umap",label=T)
c0_markers <- FindMarkers(cluster, ident.1 = 0, logfc.threshold = 0.25, test.use = "roc", only.pos = TRUE)
c1_markers <- FindMarkers(cluster, ident.1 = 1, logfc.threshold = 0.25, test.use = "roc", only.pos = TRUE)
```


# Use subclusters for annotations in the original data 
```{r}
data$subcluster = as.character(Idents(data))
data$subcluster[Cells(cluster)] = paste("CP",Idents(cluster)) # Change old cluster name from CP to your desired cluster name 
Idents(data) = data$subcluster
DimPlot(data,group.by="subcluster",reduction="umap",label=T)
```

# Rename clusters  
```{r}
data <- RenameIdents(data = data, 'cluster_old' = "cluster_new") # rename cluster_old to cluster_new
```


## Combine different objects
```{r}
combined <- merge(ctr_D35, y = list(ctr_D56, ctr_D97, D35_BMP, D56_BMP, D97_BMP, D35_BMP, D56_BMP, D97_BMP), project = "CombinedProject")
```


```{r}
combined <- NormalizeData(combined)
combined <- FindVariableFeatures(combined, selection.method = "vst", nfeatures = 2000)
combined <- ScaleData(combined)
combined <- RunPCA(combined, features = VariableFeatures(object = combined))
combined <- FindNeighbors(combined, dims = 1:15)
combined <- RunUMAP(combined, dims = 1:15)
DimPlot(combined, reduction = "umap")
```



# Go terms enrichment analysis
```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(DOSE)
```

```{r}
genes = c("gene1","gene2") # Change into the genes of your interest 

ego <- enrichGO(gene         = genes,
                OrgDb        = org.Hs.eg.db,
                keyType      = "SYMBOL",      
                ont          = "ALL",         # Ontology: BP, CC, MF, or ALL
                pAdjustMethod = "BH",         # p-value adjustment method
                pvalueCutoff  = 0.05,         # p-value cutoff
                qvalueCutoff  = 0.05,         # q-value cutoff
                readable      = FALSE)
# View results
head(ego)
dotplot(ego, showCategory = 10, orderBy = "GeneRatio") + ggtitle("GO Enrichment Dotplot")
```






